---
title: "EDS Assignment 7"
subtitle: "Review of SS3sim"
author: "Mico Kinneen"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  pdf_document:
    latex_engine : pdflatex
  toc: yes
  toc-depth: 3
urlcolor: blue
bibliography: refs.bib
csl: fisheries_research.csl
link-citations: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tinytex)
library(knitcitations)
```


# Package title

`{SS3sim}`




# Location

The package can be downloaded from Cran or from Github.
1) [Cran](https://cran.r-project.org/web/packages/SS3seem/index.html)  

```{r eval = FALSE}
install.packages('ss3sim', 
                 repos='https://cran.r-project.org/web/packages/ss3sim/index.html'
                 , dependencies=TRUE)
```
  
2) Github

```{r eval = FALSE}
remotes::install_github(
  "ss3sim/ss3sim",
  ref = "main",
  build_vignettes = TRUE, dependencies = TRUE
)
```



# Vignette(s)

`{ss3sim}` has a number of useful help files.

There are three documents which are useful when working with `{ss3sim}`. The first
is the .README, which is comprehensive and offers notes on the installation,
file structure and an example simulation of the package. The README.md can be found
[here](https://github.com/ss3sim/ss3sim/blob/main/README.md)

There are three vignettes available on the github which cover  
   1) [Introduction](https://ss3sim.github.io/ss3sim/articles/introduction.html)  
   2) [Making models](https://github.com/ss3sim/ss3sim/blob/main/vignettes/making-models.Rmd)  
   3) [Modifying models](https://github.com/ss3sim/ss3sim/blob/main/vignettes/modifying-models.Rmd).  

The package is actively under development by NOAA, and as such the 
[discussion board](https://github.com/ss3sim/ss3sim/discussions),on github is particularly useful for resolving queries.

There is also an introduction [avilable on CRAN](https://cran.r-project.org/web/packages/ss3sim/vignettes/introduction.html)
however it is deprecated and following the instructions will not allow the user
to create simulations. This is particularly problematic as when Google is used
to [search for introductory material](https://www.google.com/search?q=introduction+to+ss3sim&oq=introduction+to+ss3sim&aqs=chrome..69i57j69i60l2.4403j0j7&sourceid=chrome&ie=UTF-8), the CRAN intro is the first result.

\pagebreak


# Application(s)

The primary application of the package is that it allows users to crate simulated
fish populations and estimate them using the [Stock Synthesis 3](https://nmfs-stock-synthesis.github.io/doc/SS330_User_Manual.html) modelling 
framework.The package has been employed for simulation testing documented in 
peer-review literature. Notable publications include:  
  
@monnahan_effect_2016  
@ono_importance_2015  
@rudd_catch_2021  



# Review

`{ss3sim}` is an r package which allows users to simulate fish stock populations
and asses them using [Stock Synthesis 3](https://nmfs-stock-synthesis.github.io/doc/SS330_User_Manual.html)
modelling framework. The package was initially developed as part of a graduate class project in 
 the school of Aquatic and fisheries Science , University of Washington, with the 
 project being led by Sean Anderson (@anderson_ss3sim_2014). Each simulation has
 three components an Operating model
(OM) which represent the true population  such as natural mortality _M_, 
recruitment _R_ and fishing mortality _F_,the Estimation method (EM) 
(Stock assessments, survey data etc) which denotes how the population is to be 
assessed and the control file, which allows the users how and when they would 
like to alter the EM and OM.

While the file structure of the does appear cumbersome at first, running a 
pre-specification (or boiler plate) is realities straightforward by following
the introduction vignette.
The package allows users to alter SS3 assessments in a relatively straightforward
fashion,compared to editing the large .text files which SS3 assessments use. 
The other nice aspect of the package is there are three boiler plates, each 
mimicking a different 
fish life-history pattern. The species are Cod _Gadus morhua_ , pacific anchovy 
_Engraulis mordax_ and yellowtail founder _Pleuronectes ferruginea_ .

I would recommend the package to anyone interested in fisheries simulation 
testing or with `{ss3sim}`. As a user who had no SS3 experience, I was able to 
alter the  boiler plates and simulate the effect of having _F_ set far higher 
than is suitable. The effect of this on a fishery  over a time period 
( I chose 100 years) could then be observed.
\pagebreak

# References